The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2011-11-16 smurf_paper.tex
Add description of COM and GAI

We estimate the common-mode, remove it from the time series, use it to
flag bad bits of data, check for time-varying bolometer
responsivities... what doesn't it do?! Also mention the optional
presence of GAI.

>>>>

﻿Fig.~\ref{fig:com} shows the time series from a selection of typical
bolometers.\footnote{The data have been flat-fielded and each time series
has been adjusted to a mean value of zero.} The similarity between most
bolometers is evident, and forms the common-mode signal - assumed to be a
consequence of variations in the atmospheric emission and fridge
temperature. This common-mode usually dominates the astronomical signal
for all but the brightest sources, and swamps faint extended structure.
The purpose of the COM and GAI models is to remove this common-mode
signal.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com.pdf}
\caption{A selection of typical bolometer time series after flatfielding
and removal of a constant baseline. It can be seen that most bolometers
exhibit a common time variation overlayed with other features.}
\label{fig:com}
\end{figure}

The COM model is the common-mode signal itself. It is a single time
series that is estimated by finding the mean of all bolometers at each
time step. Bolometer values that have been flagged as unusable for any
reason are excluded from the mean.

Since bolometers have varying sensitivities, the amplitude of the
common-mode variations will also vary from bolometer to bolometer. So
comparing each bolometer time stream with the common-mode signal allows
an estimate of the bolometer sensitivity to be obtained. In practice, a
least squares linear fit is performed between the bolometer time series
and the common-mode to determine a gain and offset for each bolometer.
The gain and offset for each bolometer is known as the GAI model. Each
bolometer value can then optionally be scaled and shifted using these
values so that all bolometers share a common (but as yet unknown)
calibration. This provides an alternative( or additional) flatfielding
strategy to that described in in section~\ref{sec:bolos}.

An option exists to cater for time-varying sensitivities. If used, the
least squares fits described above are performed on short blocks of
contiguous time slices, providing multiple gain and offset values for
each bolometer (one pair for each block of time slices). The gain and
offset at any required time slice can then be found by interpolation
between these values.

It can be seen from Fig.~\ref{fig:com} that some bolometers depart
radically from the common-mode, indicating some problem within the
bolometer. Such bolometers are identified by calculating the Pearson
coefficient of correlation between each bolometer time stream and the
common mode. Bolometers for which the correlation coefficient is below a
specified limit, or which have a negative gain, are flagged as bad in
order to omit them from the final map. If the option described above for
handling time-varying sensitivities is used, then a correlation
coefficient can be determined for each individual block of time slices.
This allows individual bad blocks to be rejected from a bolometer time
stream, rather than rejecting the whole bolometer. This is beneficial for
bolometers that follow the common-mode for most of the time but suffer
from transient problems at specific times.

The common-mode is calculated from all bolometers that have not
previously been flagged as bad. Thus aberrant bolometers with low
correlation coefficients will be included, polluting the common-mode. For
this reason, an iterative approach is taken to estimating the common-mode.
The initial common-mode estimate is formed from all remaining bolometers.
Aberrant bolometers with low correlation coefficients are then identified
by comparing each bolometer time stream with the common-mode. Such
aberrant bolometers are flagged as bad, and a new estimate of the
common-mode is produced omitting the newly flagged bolometers. This
process is repeated until no further bolometers are rejected.

Any astronomical sources that are smaller than the array size will
contribute signal to some bolometers but not other bolometers, thus
biassing the simple mean used to estimate the common mode. However, on
each iteration of the map-making algorithm illustrated in
Fig.~\ref{fig:dimm}, a large fraction of the remaining astronomical
signal is extracted from the bolometer time series and transferred to the
output map, resulting in subsequent estimates of the common-mode being
more accurate.

Any extended astronomical emission on a scale comparable to or larger
than the spatial extent of the area used to estimate the common-mode will
contribute a similar signal to all bolometers. Therefore such extended
emission is indistinguishable from the other sources of common-mode
signal (fridge and atmosphere variations) and will be removed by the COM
model. This places a limit on the spatial extent of astronomical
structure that can be recovered.

For this reason, the usual practice is to estimate a single COM model by
examining data from all four sub-arrays in each waveband, since this
allow spatial structure on the scale of the whole focal plane to be
recovered. However, sometimes there is evidence that the common-mode
differs from one array to another, and so an option exists to estimate a
separate COM model for each individual subarray, with a consequent
lowering in the scale of spatial structure that can be recovered.
